%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% AI Hub Application Resume
% LaTeX Template for Piter Garcia
% Tailored for RIT AI Hub Internship
% Updated March 2026
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[10pt]{article}
\usepackage[letterpaper,margin=0.38in]{geometry}
\usepackage{fontawesome5}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage[hidelinks]{hyperref}
\usepackage{fancyhdr}

% Color definitions
\definecolor{accent}{RGB}{0,100,200}
\definecolor{darkgray}{RGB}{64,64,64}

% Section formatting
\titleformat{\section}{\large\bfseries\color{accent}}{}{0em}{}[\titlerule]
\titleformat{\subsection}{\normalsize\bfseries}{}{0em}{}

% Header setup
\pagestyle{fancy}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt}

% List spacing
\setlist[itemize]{itemsep=1.5pt, topsep=2pt}

% Custom commands
\newcommand{\resumeItem}[1]{\item\small{#1}}
\newcommand{\resumeSubheading}[4]{
  \vspace{4pt}
  \noindent
  \begin{tabular*}{\textwidth}[t]{@{}l@{\extracolsep{\fill}}r@{}}
    \textbf{#1} & #2 \\
    \textit{\small #3} & \textit{\small #4} \\
  \end{tabular*}\vspace{-4pt}
}

\begin{document}

%-------------------------------------------
% HEADER
%-------------------------------------------

\begin{center}
    {\Huge \textbf{Piter Z. Garcia Bautista}} \\
    \vspace{4pt}
    {\large Graduate Assistant in Quantum \& AI Research | Applied AI, Data Science, Reproducible Systems} \\
    \vspace{4pt}
    \faMapMarker* \, Rochester, NY \quad
    \faPhone \, +1 (646) 288-3300 \quad
    \faEnvelope \, \href{mailto:pzg8794@rit.edu}{pzg8794@rit.edu} \\
    \faLinkedin \, \href{https://linkedin.com/in/garciapiterz}{linkedin.com/in/garciapiterz} \quad
    \faGithub \, \href{https://github.com/pzg8794}{github.com/pzg8794} \quad
    \faGlobe \, \href{https://garciapiterz.com}{garciapiterz.com}
\end{center}

%-------------------------------------------
% SUMMARY
%-------------------------------------------

\section{Summary}
\vspace{-4pt}
Graduate researcher with faculty-facing experience in quantum and AI research, applied data science, healthcare analytics, and reproducible evaluation. Current work focuses on threat-aware quantum routing, qubit allocation, and research infrastructure, while the broader research direction connects fairness, equity, and performance in quantum and clinical settings. Brings prior industry experience building Python- and pandas-based analytics workflows in healthcare and production environments, with strong experience in data cleaning, exploratory analysis, supervised machine learning, and reproducible reporting.

%-------------------------------------------
% TECHNICAL SKILLS
%-------------------------------------------

\section{Technical Skills}

\textbf{Programming:} Python, R, Java, C++, C\#, SQL, MATLAB, JavaScript, Bash \\
\textbf{Applied AI / ML:} scikit-learn, TensorFlow, Keras, PyTorch, XGBoost, supervised learning, feature engineering, model evaluation, fairness-aware ML \\
\textbf{Data Science:} pandas, NumPy, SciPy, Matplotlib, Plotly, Seaborn, data cleaning, EDA, statistical analysis, benchmarking \\
\textbf{Cloud / Reproducibility:} AWS (\textit{primary industry cloud}), GCP (\textit{research workflows}), Docker, Git/GitHub, Linux/Unix \\
\textbf{Computational Research:} bioinformatics pipelines, RNA structure modeling, reproducible experiments, technical reporting

%-------------------------------------------
% RELEVANT EXPERIENCE
%-------------------------------------------

\section{Relevant Experience}

\resumeSubheading
{Graduate Assistant -- Quantum \& AI Research}{Fall 2025 -- Present}
{Software Engineering Department, Rochester Institute of Technology}{Rochester, NY}
\begin{itemize}[leftmargin=1.2em]
\resumeItem{Built a Python-based \textbf{multi-testbed experimental framework} to support \textbf{threat-aware quantum routing}, online path selection, and qubit allocation using \textbf{multi-armed bandit (MAB)}, contextual bandit, and neural decision methods.}
\resumeItem{Build and maintain comparison-ready evaluation workflows across stochastic and adversarial threat environments for benchmark quantum-network routing and allocation experiments.}
\resumeItem{Develop rerunnable workflows across \textbf{Colab, local, and GCP} environments with a shared research hub built through Colab, GCP-backed pipelines, automated validation, structured logging, shared datasets, and reproducible experiment tracking.}
\resumeItem{Evaluate AI-driven routing methods through large-scale experiments, generating benchmark analyses, technical reports, and manuscript-ready research artifacts for faculty-led publications and follow-on research.}
\end{itemize}

\resumeSubheading
{Teaching Placement -- Computer Science and Inclusion}{2025 -- Present}
{Pine Brook Elementary School}{Rochester, NY}
\begin{itemize}[leftmargin=1.2em]
\resumeItem{Support elementary computer science instruction in Pine Brook's IGNITE program through observation, assistantship, co-teaching, and gradually increasing lesson leadership.}
\resumeItem{Co-design and deliver standards-aligned lessons integrating coding, robotics, digital fluency, and responsible AI concepts through hands-on, project-based learning for elementary students.}
\resumeItem{Use differentiated materials, formative checks, and UDL-informed supports to strengthen accessibility and participation, while translating complex technical ideas into clear, usable workflows for students and collaborators.}
\end{itemize}

\resumeSubheading
{Lead Researcher -- EQUITAS and the Puzzle Plan}{2024 -- Present}
{Independent Research Initiative, RIT \& University of Rochester}{Rochester, NY}
\begin{itemize}[leftmargin=1.2em]
\resumeItem{Develop a cross-institutional research program linking healthcare analytics, RNA bioinformatics, fairness-aware diagnostics, bandit-based decision systems, disability-informed governance, and inclusive education.}
\resumeItem{Build the technical backbone of this work through hospital readmission prediction, RNA secondary structure modeling, fairness-aware bioinformatics, and diagnostic AI governance, always linking predictive performance to equitable outcomes.}
\resumeItem{Operationalize fairness, equity, and performance as joint design goals through fairness audits, mitigation strategies, governance narratives, workshop design, and community-informed research documentation.}
\end{itemize}

\resumeSubheading
{Data Solutions Engineer}{Sep 2022 -- Aug 2024}
{VEDADATA Inc.}{New York, NY}
\begin{itemize}[leftmargin=1.2em]
\resumeItem{Designed and implemented \textbf{Python-based data pipelines} and analytics workflows for business applications, emphasizing large-scale ingestion, cleaning, preprocessing, and validation.}
\resumeItem{Built maintainable cloud workflows on \textbf{AWS}, supporting reproducible data transformation, and downstream analytics readiness.}
\resumeItem{Applied quality checks, statistical analysis, and practices to improve traceability, reliability, and usability of production data systems.}
\end{itemize}

\resumeSubheading
{AI Data Solutions Engineer}{Jan 2021 -- Sep 2022}
{VIOME Inc.}{New York, NY}
\begin{itemize}[leftmargin=1.2em]
\resumeItem{Developed \textbf{applied AI and machine learning workflows} in Python for healthcare use cases, including preprocessing, feature engineering, model experimentation, and evaluation.}
\resumeItem{Implemented supervised ML workflows using \textbf{scikit-learn} and related tools to support predictive and recommendation-oriented health applications.}
\resumeItem{Worked extensively in \textbf{AWS}-based environments to support reproducible model development, deployment-oriented data preparation, and reliable evaluation pipelines.}
\end{itemize}

%-------------------------------------------
% EDUCATION
%-------------------------------------------

\section{Education}

\resumeSubheading
{Master of Science in Data Science}{Expected Aug 2026}
{Rochester Institute of Technology}{Rochester, NY}
\begin{itemize}[leftmargin=1.2em]
\resumeItem{GPA: 3.9+ \;|\; Research focus: Fairness-Aware AI in Quantum \& Clinical Settings}
\resumeItem{Relevant Coursework: Data Science Foundations, Software Engineering for Data Science, Bioinformatics Algorithms, Data-Driven Knowledge Discovery, Neural Networks, Applied Statistics}
\end{itemize}

\resumeSubheading
{Master of Science in Teaching Computer Science K--12}{Expected Aug 2026}
{University of Rochester, Warner School of Education}{Rochester, NY}
\begin{itemize}[leftmargin=1.2em]
\resumeItem{GPA: 3.9+ \;|\; NSF Robert Noyce Scholar \;|\; Advanced Certificate in Leadership in Disability and Inclusive Practices}
\resumeItem{Relevant Coursework: Disability and Schools, Assessment in Instructional Contexts, Inclusive Classrooms, CS Standards Integration}
\end{itemize}

\resumeSubheading
{Master of Science in Computer Science}{2015}
{Rochester Institute of Technology}{Rochester, NY}
\begin{itemize}[leftmargin=1.2em]
\resumeItem{Concentrations: Artificial Intelligence, Big Data Analytics, Computer Graphics}
\end{itemize}

\resumeSubheading
{Bachelor of Science in Computer Engineering}{2012}
{Farmingdale State College}{Farmingdale, NY}
\begin{itemize}[leftmargin=1.2em]
\resumeItem{Dual Degree: Electrical Engineering \& Computer Engineering Technology}
\end{itemize}

%-------------------------------------------
% RELEVANT AWARDS & RECOGNITION
%-------------------------------------------

\section{Relevant Awards \& Recognition}

\begin{itemize}[leftmargin=*]
\item \small{\textbf{NSF Robert Noyce Scholar} (2024--2026)}
\item \small{\textbf{MS International Scholarship} (2012--2015) -- merit-based graduate funding through RIT and MESCYT}
\item \small{\textbf{Honorable Mention Technology Award} (2012) -- NYS STEP/CSTEP capstone recognition}
\item \small{\textbf{IEEE Alumni Member} (2009--Present)}
\end{itemize}

%-------------------------------------------
% FACULTY RESEARCH ADVISERS / COLLABORATORS
%-------------------------------------------

\section{Research Advisers and Collaborators}

Active research advisers and collaborators connected to my work.

\begin{itemize}[leftmargin=*]
\item \small{\textbf{Dr. Daniel Krutz and Dr. Travis Desell} -- Faculty collaboration and supervision in the current quantum and AI graduate assistantship work, including experiment infrastructure, validation workflows, paper development, quantum routing, and multi-testbed research systems}
\item \small{\textbf{Dr. Qi Yu} -- Guidance in fairness-aware machine learning, data-driven knowledge discovery, and the development of equitable bioinformatics work}
\item \small{\textbf{Dr. Fernando Rodriguez} -- Bioinformatics and computational biology guidance for RNA secondary structure modeling, biomarker analysis, and diagnostic decision-support research}
\end{itemize}

\vspace{8pt}
\small{Formal references, portfolio materials, code samples, and research papers are available upon request.}

\end{document}
